Computational musicology, ????, Profit

### 2016-07-24 paper.tex
Incremental edit.

Amnon Eden classifies the way that computer scientists think about their craft into rationalist, technocratic and scientific paradigms\cite{Three Paradigms}.
These conceptual models position programmers as mathematicians, engineers or scientists respectively. They vary in how they
produce knowledge, but all three belong to Snow's scientific culture. None of them value or use literary methods.

To properly understand computational musicology, let us contrast with computational physics. Computational physics papers cannot
influence natural phenomena because physics cannot influence natural phenomena.
There is an inviolable barrier between the subject (physics) and
the object (natural phenomena), so the field of study cannot contribute to its object. Planets orbited the sun long before
Copernicus presented a mathematical model of heliocentrism. Kepler's elliptical orbits did not consist of a new and improved way
for celestial bodies to move. Even Galileo's emperical observations were not able to provoke reform of the solar system itself.

@@

Marvin Minksy recognises this problem in an essay called `Music, Mind, and Meaning'. He suggests that computational
musicologists ``would like to study Mozart's music the way scientists analyze the spectrum of a distant star", but that music
is not the kind of object that can be studied ``uninfluenced by custom or constraint of commerce"\cite{Music Mind and Meaning}.
I deliberately refer to Minsky's work as an essay and not as a paper. It does not conform to the conventions of a scientific
or mathematical paper. He uses the first person. He speculates. He draws connections that he cannot mathematically formalise.
He makes painstaking reference to the details of a specific artistic work (Beethoven's Fifth). He states no theorems.

@@

\section{Air on the {\textbackslash}G String}
To view music within Eden's scientific paradigm, we might begin by understanding music as data.
Very great quantities of data can exhibit very little complexity. Consider the following Clojure expression, which
evaluates to a sequence of $10^{15}$ uppercase G characters:

@@

Through his Library of Babel, Borges is playing with our notions of invention versus discovery. Mathematicians and computer
scientists struggle with whether they are the inventors of new theories or discoverers. Philip Wadler, for example,
argues that the lambda calculus is so fundamental that it deserves to be considered a universal concept that
was discovered\cite{Propositions as Types}. By positing a library where any work may be found if you know where to look, Borges
suggests that even literary authorship could be considered a search problem and therefore a matter of discovery rather than
creation.

@@

Computational musicology is part of a rich tradition of information theoretical art and criticism. Music's deep structure and
cultural significance give computational musicology great potential as a source of artistic analysis and inspiration.
This paper represents a self-evidencing claim that computational musicology can contribute to musical ideas and appreciation.

>>>>

Amnon Eden classifies the way that computer scientists think about their craft into rationalist, technocratic and scientific paradigms\cite{Three Paradigms}.
These conceptual models position programmers as mathematicians, engineers or scientists respectively. They vary in how they
produce knowledge, but all three belong to Snow's scientific culture. None value or use literary methods.

To properly understand computational musicology, let us start with an example of computational methods that do not mix Snow's
cultural categories. Computational physics papers cannot
influence natural phenomena because physics cannot influence natural phenomena.
There is an inviolable barrier between the subject (physics) and
the object (natural phenomena), so the field of study cannot interact with its object.
Kepler's elliptical orbits improved our understanding of the motions of celestial bodies, but the planets themselves
remained unaffected.

@@

Marvin Minksy recognises this problem in an essay called `Music, Mind and Meaning'. He suggests that computational
musicologists ``would like to study Mozart's music the way scientists analyze the spectrum of a distant star", but that music
is not the kind of object that can be studied ``uninfluenced by custom or constraint of commerce"\cite{Music Mind and Meaning}.

I deliberately refer to Minsky's work as an essay and not as a paper. It does not conform to the conventions of a scientific
or mathematical paper. He uses the first person. He speculates. He draws connections that he cannot mathematically formalise.
He engages deeply with the detail of an individual artistic work (Beethoven's fifth symphony). He states no theorems, though he
does propose a model of musical perception. A conventional scientific paper would not have allowed him to express the same breadth 
of insight.

@@

\section{Air on the {\textbackslash}G String}
To return to Eden's scientific paradigm for a moment, we might begin by understanding music as data.
Very great quantities of data can exhibit very little complexity. Consider the following Clojure expression, which
evaluates to a sequence of $10^{15}$ uppercase G characters:

@@

Through the Library of Babel, Borges plays with our notions of invention versus discovery. Mathematicians and computer
scientists struggle with whether they are the inventors of new theories or discoverers. Philip Wadler, for example,
argues that the lambda calculus is so fundamental that it deserves to be considered a universal concept that
was discovered\cite{Propositions as Types}. By positing a library where any work may be found if you know where to look, Borges
suggests that even literary authorship could be considered a search problem and therefore a matter of discovery rather than
creation.

@@

Computational musicology is part of a rich tradition of information theoretical art and criticism. Music's deep structure and
cultural significance give computational musicology great potential as a source of artistic analysis and inspiration.
This paper represents a self-supporting claim that computational musicology can contribute to musical ideas and appreciation as
a form of creative criticism.
